---
title: "Laboratory 5"
author: "Clay Curry"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    fig_caption: true
    toc_depth: 2
    highlights: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

Getting Working Directory:
```{r}
getwd()
```

# Task 2

## Subtask 3

### simulate a binomial experiment where n=10,p=0.7, and Y=number of successes.


```{r, cached=TRUE}
## sampling function
library(MATH4753CURR0011)
mybin(n = 10, iter = 10, p = .7)
```
10 iterations

## Subtask 4

### Record the plots for each of the following (Use color)


```{r}
## sampling function
mybin(n = 10, iter = 100, p = .7)
```
100 iterations

```{r}
## sampling function
mybin(n = 10, iter = 200, p = .7)
```
200 iterations

```{r}
## sampling function
mybin(n = 10, iter = 500, p = .7)
```
500 iterations

```{r}
## sampling function
mybin(n = 10, iter = 1000, p = .7)
```
1000 iterations

```{r}
## sampling function
mybin(n = 10, iter = 10000, p = .7)
```
10,000 iterations

## Subtask 5

### Verify that each value in the table is approximately correct using dbinom()
```{r, cached=TRUE, collapse=FALSE}
bin = dbinom(x = 0:10, size = 10, prob = 0.7)
sim = mybin(iter = 10000, n = 10, p = 0.7, mkplot = FALSE)
sim = with(data.frame(sim), Freq)

sd(bin - sim)
```
Notice in the above code that the standard deviation of 10,000 iterations from the **expected value** of the binomial distribution is `r round(sd(bin - sim),4)`, essentially 0.

# Task 3
Suppose that there is a bag of 20 marbles, 12 white "1" and 8 black "0".
Using the sample() function, create a sample of size n=5 without replacement.
```{r}
sample(x = ifelse(1:20 <= 8, 0, 1), size = 5, replace = TRUE)
```

```{r}
sample(x = ifelse(1:20 <= 8, 0, 1), size = 5, replace = TRUE)
```

```{r}
myhyper=function(iter=100,N=20,r=12,n=5){
# make a matrix to hold the samples
#initially filled with NA's
sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)
#Make a vector to hold the number of successes over the trials
succ=c()
for( i in 1:iter){
#Fill each column with a new sample
sam.mat[,i]=sample(rep(c(1,0),c(r,N-r)),n,replace=FALSE)
#Calculate a statistic from the sample (this case it is the sum)
succ[i]=sum(sam.mat[,i])
}
#Make a table of successes
succ.tab=table(factor(succ,levels=0:n))
#Make a barplot of the proportions
barplot(succ.tab/(iter), col=rainbow(n+1), main="HYPERGEOMETRIC simulation", xlab="Number of successes")
succ.tab/iter
}
myhyper(iter=10000,n=19, N=20,r=12)

dhyper(x=0:19, m=12, n=8, k=19)
```

# Task 4

```{r}
mysample=function(n=1000, iter=1,time=1){
for( i in 1:iter){
#make a sample
s=sample(1:10,n,replace=TRUE)
# turn the sample into a factor
sf=factor(s,levels=1:10)
#make a barplot
barplot(table(sf)/n,beside=TRUE,col=rainbow(10), 
main=paste("Example sample()", " iteration ", i, " n= ", n,sep="") ,
ylim=c(0,0.2)
)
#release the table
Sys.sleep(time)
}
}

mysample()
```

# Task 5
```{r}
choose(8,4)

1-ppois(q = 4,lambda = 2)
  
dnbinom(x=10-3,size = 3, prob = 0.4)

pbinom(q=8,size=15,prob=0.4)
```


# Task 6

```{r}
getSample(num0s = 10, num1s = 15, sampleSize = 2, replace = TRUE)
```

