---
title: "Laboratory 5"
author: "Clay Curry"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    fig_caption: true
    toc_depth: 2
    highlights: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

Getting Working Directory:
```{r}
getwd()
```

# Task 2

## Subtask 3

### simulate a binomial experiment where n=10,p=0.7, and Y=number of successes.


```{r, cached=TRUE}
## sampling function
library(MATH4753CURR0011)
mybin(n = 10, iter = 10, p = .7)
```
10 iterations

## Subtask 4

### Record the plots for each of the following (Use color)


```{r}
## sampling function
mybin(n = 10, iter = 100, p = .7)
```
100 iterations

```{r}
## sampling function
mybin(n = 10, iter = 200, p = .7)
```
200 iterations

```{r}
## sampling function
mybin(n = 10, iter = 500, p = .7)
```
500 iterations

```{r}
## sampling function
mybin(n = 10, iter = 1000, p = .7)
```
1000 iterations

```{r}
## sampling function
mybin(n = 10, iter = 10000, p = .7)
```
10,000 iterations

## Subtask 5

### Verify that each value in the table is approximately correct using dbinom()
```{r, cached=TRUE, collapse=FALSE}
bin = dbinom(x = 0:10, size = 10, prob = 0.7)
sim = mybin(iter = 10000, n = 10, p = 0.7, mkplot = FALSE)
sim = with(data.frame(sim), Freq)

sd(bin - sim)
```
Notice in the above code that the standard deviation of 10,000 iterations from the **expected value** of the binomial distribution is `r round(sd(bin - sim),4)`, essentially 0.

