---
title: "lab6"
author: "Clay Curry"
date: "10/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1
Getting the directory
```{r}
getwd()
```

# Task 2
Creating 4 plots with bounds at .001 < q < .999
```{r}
layout(matrix(1:4, nr=2,nc=2))

curve(dnorm(x, mean=10,sd=4),
      xlim=c(qnorm(.001, mean = 10, sd = 4),
             qnorm(.999, mean = 10, sd = 4)))

curve(dnorm(x, mean=10,sd=2),
      xlim=c(qnorm(.001, mean = 10, sd = 2),
             qnorm(.999, mean = 10, sd = 2)))

curve(dnorm(x, mean=5,sd=10),
      xlim=c(qnorm(.001, mean = 5, sd = 10),
             qnorm(.999, mean = 5, sd = 10)))

curve(dnorm(x, mean=5,sd=1/2),      
      xlim=c(qnorm(.001, mean = 5, sd = 1/2),
             qnorm(.999, mean = 5, sd = 1/2)))

```

---

Now things get interesting

```{r cached=TRUE}
library(ggplot2)
```


---

plotting	Y∼N(0,1),P(Y≥2)

```{r}
normargs <- list(mean = 0, sd = 1)

n <- ggplot(data.frame(x = c(-2,4)), aes(x)) + 
  stat_function(fun = dnorm, args = normargs , geom = "area", col = "black") 
  

n <- n + stat_function(fun = dnorm, args = normargs, geom = "area", fill = "green", alpha = 2,xlim = c(2,4)) 

n <- n + xlab("X") + ylab("Density")
n
```


---

plotting	Y∼N(μ=4,σ=2),P(1≤Y<5)

```{r}
normargs <- list(mean = 4, sd = 2)

n <- ggplot(data.frame(x = c(-2,8)), aes(x)) + 
  stat_function(fun = dnorm, args = normargs , geom = "area", col = "black") 
  

n <- n + stat_function(fun = dnorm, args = normargs, geom = "area", fill = "green", alpha = 2,xlim = c(1,5)) 

n <- n + xlab("X") + ylab("Density")
n
```

---

plotting	Y∼N(μ=10,σ=4),P(Y<10)

```{r}
normargs <- list(mean = 10, sd = 4)

n <- ggplot(data.frame(x = c(-2,14)), aes(x)) + 
  stat_function(fun = dnorm, args = normargs , geom = "area", col = "black") 
  

n <- n + stat_function(fun = dnorm, args = normargs, geom = "area", fill = "green", alpha = 2,xlim = c(-2,10)) 

n <- n + xlab("X") + ylab("Density")
n
```

---

plotting	Y∼N(μ=-2,σ=1/2),P(-3<Y≤ -2)
```{r}
normargs <- list(mean = -2, sd = 1/2)

n <- ggplot(data.frame(x = c(-4,0.5)), aes(x)) + 
  stat_function(fun = dnorm, args = normargs , geom = "area", col = "black") 
  

n <- n + stat_function(fun = dnorm, args = normargs, geom = "area", fill = "green", alpha = 2,xlim = c(-3,-2)) 

n <- n + xlab("X") + ylab("Density")
n
```

# Task 3

	We will now study the gamma distribution. This has two parameters, shape (>0) and scale (>0). You will use dgamma() and pgamma()
	Notice that the random variable Y takes only positive values.
	On the one plotting surface overlay the following three plots (NB – Don’t use layout)
	
---
	
Y∼Gamma(shape=1,scale=1)

Y∼Gamma(shape=3,scale=1)

Y∼Gamma(shape=5,scale=1)


```{r}
curve(dgamma(x,shape=1,scale=1),xlim=c(0,10),ylim=c(0,1),col="Red",lwd=2,
ylab="Gamma density", main="Beta=1")
text(x=1,y=0.8,paste("alpha=",1))
curve(dgamma(x,shape=3,scale=1),xlim=c(0,10),ylim=c(0,1),add=TRUE,lwd=2)
text(x=2.2,y=0.3, paste("alpha=",3))
curve(dgamma(x,shape=5,scale=1),xlim=c(0,10),ylim=c(0,1),add=TRUE, col="Blue",lwd=2)
text(x=6,y=0.2,paste("alpha=",5))
```


Plot the following regions and probabilities in R
	Y~Gamma(shape=3,scale=2),P(2<Y<5)
	Y∼Gamma(shape=6,scale=3),P(1≤Y≤4)
	Y∼Gamma(shape=2,scale=4),P(3≤Y<6)
