---
title: "Assignment 1"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
A table depicting grade calculation for this class, including grading scale.
```{r}
GradCalc <- data.frame(c(15, 10, 10, 10, 5, 20, 30))
GradCalc <- data.frame(GradCalc, c("Assignments", "Laboratories", "Projects (1/3 Project 1; 2/3 Project 2)", "In class Quizzes", "Canvas chapter quizzes", "Mid-Term Exams (x2)", "Final"))
GradCalc <- data.frame(GradCalc, c("A", "B", "C", "D", "F", "", ""))
GradCalc <- data.frame(GradCalc, c("90s","80s", "60s and 70s", "50s", "<50", "",""))

knitr::kable(GradCalc, col.names = c("Weight %", "Component", "Letter", "Score"))
```

## Question 2


### 2A.


```{r}
DDT<-read.csv("../../MyData/DDT.csv") # Reads in DDT
m=with(DDT, as.numeric(factor(MILE))) # A, could also be m=as.numeric(DDT$MILE)
n<-length(unique(m)) #B
mileCol <- with(DDT, m, col = 1:n)
coplot(LENGTH~WEIGHT|RIVER*SPECIES, data = DDT, col = mileCol)

```


###     2B.
  The lower left 3 plots show the length of catfish vs weight across the rivers FCM, LCM, SCM, TRM.
  
###     2C.
  The factor() function extracts the discrete values from DDT$MILE. The as.numeric() function converts factor() into a numerical vector, and with() evaluates as.numeric(factor()) in a local environment constructed from DDT.
  
###     2D.
  The unique() function returns a vector containing a single instance of each mile observation. The length() function gets the length of this vector.
  
###     2E.
  Plots are empty where the frequency of observations equals zero. Notice how LMBASS and SMBUFFALO were not observed in any river outside of TRM.

```{r}
DDT[DDT$RIVER!="TRM" & (DDT$SPECIES=="SMBUFFALO" | DDT$SPECIES=="LMBASS"),]
```

