---
title: "Assignment 1"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
A table depicting grade calculation for this class, including grading scale.
```{r}
GradCalc <- data.frame(c(15, 10, 10, 10, 5, 20, 30))
GradCalc <- data.frame(GradCalc, c("Assignments", "Laboratories", "Projects (1/3 Project 1; 2/3 Project 2)", "In class Quizzes", "Canvas chapter quizzes", "Mid-Term Exams (x2)", "Final"))
GradCalc <- data.frame(GradCalc, c("A", "B", "C", "D", "F", "", ""))
GradCalc <- data.frame(GradCalc, c("90s","80s", "60s and 70s", "50s", "<50", "",""))

knitr::kable(GradCalc, col.names = c("Weight %", "Component", "Letter", "Score"), label = "Grade Calculation")
```

## Question 2


### 2A.


```{r}
DDT<-read.csv("../../MyData/DDT.csv") # Reads in DDT
m=with(DDT, as.numeric(factor(MILE))) # A, could also be m=as.numeric(DDT$MILE)
n<-length(unique(m)) #B
mileCol <- with(DDT, m, col = 1:n)
coplot(LENGTH~WEIGHT|RIVER*SPECIES, data = DDT, col = mileCol)

```


### 2B.
The lower left 3 plots show the length of catfish vs weight across the rivers FCM, LCM, SCM, TRM.

### 2C.
The factor() function extracts the discrete values from DDT$MILE. The as.numeric() function converts factor() into a numerical vector, and with() evaluates as.numeric(factor()) in a local environment constructed from DDT.

### 2D.
The unique() function returns a vector containing a single instance of each mile observation. The length() function gets the length of this vector.

### 2E.
Plots are empty where the frequency of observations equals zero. Notice how LMBASS and SMBUFFALO were not observed in any river outside of TRM.

```{r}
DDT[DDT$RIVER!="TRM" & (DDT$SPECIES=="SMBUFFALO" | DDT$SPECIES=="LMBASS"),]
```

### 2F.
The arithmetic mean for DDT found in the sample of CCATFISH caught in the FCM river:

```{r}
mean(with(DDT[DDT$RIVER=="FCM" & DDT$SPECIES=="CCATFISH",], DDT))
```




## Question 3

### a.
ratio

### b.
ratio

### c.
nominal

### d.
ratio

### e.
ordinal

### f.
ratio

### g.
nominal

## Question 4
Let $\Omega$ denote some data set containing (possibly infinite) experimental units characterizing some phenomenon of interest to us.

### Simple Random Sampling
The set $S \subset \Omega$ constitutes a **simple random sample** when every $\omega \in \Omega$ has an equal chance of being included in $S$.

An example would be selecting data points at random from a hat while blindfolded. There are other, more sophisticated techniques.

### Stratified Random Sampling
The set $S \subset \Omega$ constitutes a **stratified random sample** when the $\Omega$ can be partitioned into two or more groups $G_1, G_2, \dots$ called *strata* where each $\omega \in G_i, \hspace{3pt}i=j$ shares some characteristic(s) with other elements in $G_i, \hspace{3pt}i=j$, which it does not share with elements in $\omega \in G_i, \hspace{3pt}i \ne j$.

An example would be a pre-election poll in which a right-leaning candidate is only interested in the ranking of important of issues among right-leaning voters, namely which issues motivate right-leaning voters to show up to the polls. This candidate may also be interested in polling left-leaning voters to gather information on which issues are likely to cause left-leaning voters show up to the polls, avoiding discussions on these issues in public settings.

### Cluster Sampling
The set $S \subset \Omega$ constitutes a **cluster sample** when the $\Omega$ can be partitioned into two or more groups $G_1, G_2, \dots$ where each $\omega \in G_i, \hspace{3pt}i=j$ constitutes a group of physically related (sequentially, geographically, ...) units.

Here are some examples: cell phone coverage in various cities, self-reported confidence levels in STEM programs according to gender, likely targets for software bugs for a program analyzed according to related modules modules.

### Systematic Sampling
The set $S \subset \Omega$ constitutes a **systematic sample** when each $\omega \in S$ is selected according to a regular rule, such as choosing at time intervals, or selecting every *k*th unit from a production. 
